---
title: "Effect of Population Density on Life Expectancy"
author: "Ojaswi Malik"
date: "`r lubridate::today()`"
output: github_document
---


## Setup

```{r setup}

library(tidyverse)
library(coefplot)
#loading gapminder() package
library(gapminder)
#loading geonames() package
library(geonames)
#loading countrycode() package
library(countrycode)

#retrieve API key stored in .Rprofile
key <- getOption("geonamesUsername")

#setting theme
theme_set(theme_minimal())
```

## Analysis

```{r building graph, message=FALSE, warning=FALSE}
#glimpsing the gapminder data
glimpse(gapminder)

# 1.loading the country information using geonames 
countryInfo <- GNcountryInfo()

# 2. using countrycode() to make country names format  same for both datasets 
countryInfo <- countryInfo %>%
  mutate(country = countrycode(sourcevar = countryInfo$countryCode, origin = "iso2c", destination = "country.name"))

# 2. Joining the 2 datsets and 3. Calculating population density
joined <- gapminder %>%
  left_join(countryInfo, by = "country" ) %>%
  #converting area into numeric to use it to calculate pop density
  mutate(areaInSqKm = as.numeric(areaInSqKm)) %>%
  mutate(pop_dens = pop/areaInSqKm) 

# 4. Plotting graph with population density
joined %>%
  ggplot(mapping = aes(x = pop_dens, y=lifeExp )) +
  geom_point(alpha = 0.1) +
  geom_smooth() + 
  scale_x_log10() +
  labs(
    title = "Relationship between Population Density and Life Expectancy",
    x = "Population Density (Population/Area in Square KM)",
    y = "Average Life Expectancy"
  )

# Comparing across continents
joined %>%
  #dropping NA values
  drop_na(continentName) %>%
  group_by(continentName) %>%
  ggplot(mapping = aes(x = pop_dens, y=lifeExp, color = continentName )) +
  geom_point(alpha = 0.1) +
  geom_smooth() + 
  scale_x_log10() +
  labs(
    title = "Relationship between Population Density and Life Expectancy",
    subtitle = "Difference across continents",
    x = "Population Density (Population/Area in Square KM)",
    y = "Average Life Expectancy",
    color = "Continent Name"
  ) + 
  theme(legend.position = "bottom")
  
```

**Observation** :

- **Overall**: As population density increases, the average life expectancy increases. Moreover, the rate of growth increases as population density nears 100.

- **Across continents**: Notably, Asia has the highest population density and follows the general trend i.e. as population density increases, the average life expectancy increases. Similiarly, Africa also has higher life expectancy when population density increases. Both Oceania and North America seem to have "W" shaped graphs, implying higher population density having initially lower high expectancy, but later there are high and low life expectancies as population density increases. Moreover, both Oceania and North America have higher average life expectancy that Asia and Africa. South America has a sharp rise in average life expectancy as population density increases. Europe initially has a negative relationship between the 2 variables, but for higher values of population density it has higher averae life expectancy.

## Coefficient Plot

This coefficient plot estimates a linear regression model for predicting Average life expectancy as a function of GDP per capital, Area in Sq KM, and population density.

```{r}
#constructing OLS model
exp_mod <- lm(lifeExp ~ pop + gdpPercap + areaInSqKm + pop_dens,
              data = joined)

#creating the plot
coefplot(exp_mod, sort = "magnitude", intercept = FALSE)

```
## Observation:
As we saw in the smoothing line above, population density has a positive relationship with average life expectancy.But it also has high variance. The other 3 coefficients: GDP per capital, Area in Sq KM, and population density also have a positive relationship respectively.

## Session Info

```{r, echo = TRUE}
devtools::session_info()
```