---
title: "Web Scraping - Internet Usage Data"
author: "Ojaswi Malik"
date: "`r lubridate::today()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rvest)
library(coefplot)

theme_set(theme_minimal())
```

## Scraping the data

```{r scraping, message=FALSE, warning=FALSE}
#reading the website
read <- read_html("http://wdi.worldbank.org/table/5.12")

#selecting the elements with css table
table <- html_nodes(read, css = "table") 

#choosing the 3rd table and renaming the headers of the columns
internet_data <- html_table(table, fill = TRUE)[[3]] %>%
  as_tibble() %>%
  rename(
    "Country" = 1,
    "2018_Internet_Usage" = 2,
    "2018_Secure_Internet_servers" = 3,
    "2017_Goods_Exports" = 4,
    "2017_Goods_Imports" = 5,
    "2017_Services_Exports" = 6
  )

#converting all data in the table to numeric so we can analyse it
new_internet_data <- internet_data %>%
  mutate(
    `2018_Internet_Usage` = as.numeric(`2018_Internet_Usage`),
    `2018_Secure_Internet_servers`= as.numeric(`2018_Secure_Internet_servers`),
    `2017_Goods_Exports` = as.numeric(`2017_Goods_Exports`),
    `2017_Goods_Imports` = as.numeric(`2017_Goods_Imports`),
    `2017_Services_Exports` = as.numeric(`2017_Services_Exports`)
  )

#deleting rows from 218 to 264 because those are not countries
new_internet_data <- new_internet_data[-c(218:264), ]

```
## Analysis

###  1. Relationship between Internet Users and Secure Internet Servers

```{r relationship1, message=FALSE, warning=FALSE}
new_internet_data %>%
  ggplot(aes(x = `2018_Internet_Usage`, y = `2018_Secure_Internet_servers`)) + 
  geom_point(alpha = 0.5) + 
  #using a smoothing line to visualise the relationship
  geom_smooth() + 
  labs(
    title = "Relationship between Internet Users and Secure Internet Servers",
    x = "% of population using the internet in 2018",
    y = "Secure internet servers per million people"
  )

```

### Observation: 
The general trend shows that as the percentage of population using the internet increases, the number of secure internet servers also increases. This makes sense as a country would need higher number of servers to support large internet user in their population.

### 2. Ranking all countries in order of Internet Users in 2018

```{r ranking, fig.height = 20}

new_internet_data %>%
  #removing NA values
  drop_na(`2018_Internet_Usage`) %>%
  #in descending order of 2018 internet usage
  ggplot(aes(x = fct_reorder(Country, `2018_Internet_Usage`), y = `2018_Internet_Usage`)) + 
  geom_col() +
  #highlighting USA in blue
  geom_col(data = new_internet_data %>%
             filter(Country == "United States"), 
           fill = "dodgerblue") +
  coord_flip() +
  labs(
    title = "Countries in order of internet users",
    subtitle = "USA highlighted",
    y = "% of population using the internet in 2018",
    x = "Country"
  )

```

### Observation:
Plotting all 216 countries, this yields a very large plot. Comparing the first and last countries there is big difference of about 98% of population using the internet. I have also highlighted the US's position in blue which is in the top 25% of all countries. In the next analyses, I will narrow down the list of countries.

### 3. Focusing on top 15 countries

```{r top15}
#getting only the top 15 countries from the above graph
top_15 <- top_n(new_internet_data, 15, `2018_Internet_Usage`)

top_15 %>%
  drop_na(`2018_Internet_Usage`) %>%
  ggplot(aes(x = fct_reorder(Country, `2018_Internet_Usage`), y = `2018_Internet_Usage`)) + 
  geom_col() +
  coord_flip() +
    labs(
    title = "Top 15 countries in order of internet users",
    y = "% of population using the internet in 2018",
    x = "Country"
  )


```

### Observation: 
These top 15 countries are really close in their percentage of population on the internet. The leading country is Qatar.

### 4. For these top 15 countries plotting good exports, imports and service exports in 2017 

```{r 2017}
#for 2017 goods exports
top_15 %>%
  drop_na(`2017_Goods_Exports`) %>%
  ggplot() +
  geom_col(aes(x = Country, y = `2017_Goods_Exports`), fill = "red") +
  coord_flip() + 
  labs(
    title = "Country wise goods exports",
    x = "Country",
    y = "% of total goods exports in 2017"
  )

#for 2017 goods imports
top_15 %>%
  drop_na(`2017_Goods_Imports`) %>%
  ggplot() +
  geom_col(aes(x = Country, y = `2017_Goods_Imports`), fill = "blue") +
  coord_flip() +
  labs(
    title = "Country wise goods imports",
    x = "Country",
    y = "% of total goods imports in 2017"
  )

#for 2017 services exports
top_15 %>%
  drop_na(`2017_Services_Exports`) %>%
  ggplot() +
  geom_col(aes(x = Country, y = `2017_Services_Exports`), fill = "purple") +
  coord_flip() +
  labs(
    title = "Country wise services exports",
    x = "Country",
    y = "% of total services exports in 2017"
  )


```


### Observation:

- **Goods, Exports**: Despite being the country with the highest % of internet users, Qatar has the lowest % of goods exports. Korea has the largest percentage of goods exported in 2017.

- **Goods, Imports**: Korea also has the largest percentage of goods imported in 2017.

- **Services, Exports**: Kuwait has the largest percentage of services exported in 2017.

### 5. Coefficient Plot
This coefficient plot estimates a linear regression model for predicting 2018 Internet Usage on average as a function of secure internet servers in 2018, goods exported and imported in 2017, and services exported in 2017.

```{r coefficient plot}
#OLS model
users_mod <- lm(`2018_Internet_Usage` ~ 
                  `2018_Secure_Internet_servers` + 
                  `2017_Goods_Exports` +
                  `2017_Goods_Imports` +
                  `2017_Services_Exports`,
                data = new_internet_data
                  )

#generating the coefficient plot
coefplot(users_mod, sort = "magnitude", intercept = FALSE)
  
```

### Observation 

The relationships highlighted from this plot:

- As good imported increases, there is a higher percentage of users on the internet on average.

- As secure internet servers increases, there are a higher % of internet users. Although, this relationship is not as strong as with goods imported.

- As services exported and goods exported increases, there is a fewer percentage of users on the internet on average.


## Session Info

```{r, echo = TRUE}
devtools::session_info()
```